---
title: "Factores maternos asociados a macrosomía neonatal: análisis descriptivo"
author: "Inés García de la Peña Marco"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 7,
  fig.height = 5
)
```  
 
# 1. Objetivo  

# 2. Cargar el dataset de análisis
```{r load-data}
df <- read_csv(
  file.path(PATH_TRANSFORMADOS, "dataset_analitico.csv")
)
```

# 3. Preparación de datos

```{r}
# Pasar el peso del recién nacido de gramos a kg
df <- df %>%
  mutate(
    peso_nacimiento = peso_nacimiento / 1000
  )
```

## 3.2. Función para establecer el estilo de la tabla  
```{r}
tabla_estilo <- function(tabla){

  tabla %>%
    add_overall(last = FALSE) %>%
    add_n() %>%
    add_p(
      test = list(
        all_continuous() ~ "wilcox.test",
        all_categorical() ~ "chisq.test"
      )
    ) %>%
    modify_fmt_fun(
      p.value ~ label_style_pvalue(digits = 3)
      )%>%
    bold_labels() %>%
    bold_p(t = 0.05) %>%
    italicize_levels()

}

tabla_estilo_strata <- function(tabla){

  tabla %>%
    add_n() %>%
    add_p(
      test = list(
        all_continuous() ~ "wilcox.test",
        all_categorical() ~ "chisq.test"
      )
    ) %>%
    modify_fmt_fun(
      p.value ~ label_style_pvalue(digits = 3)
      )%>%
    bold_labels() %>%
    bold_p(t = 0.05) %>%
    italicize_levels()

}
```  
  
## 3.3. Función para guardar las tablas  
```{r}
save_table <- function(tbl, filename){

  tbl %>%
    as_gt() %>%
    gt::gtsave(
      paste0(
        "Y:/PROYECTOS/2024 Salud perinatal (Luis-Aída-Sol)/Desarrollo/Results/tablas/",
        filename
      )
    )

}
```  
  
## 4. Tablas  
  
## 4.1. Tabla 1: Baseline characteristics according to maternal origin (Spain vs Foreign) and LGA status.

```{r}
tabla1 <- df %>%
  select(

    madre_extranjera,
    macrosomia,

    ## Maternal characteristics
    edad,
    talla,
    peso_inicial,
    imc_inicial,
    ganancia_peso,
    consumo_tabaco,
    consumo_alcohol,

    ## Pregnancy
    edad_gestacional_nacimiento,
    tipo_parto,

    ## Neonatal
    peso_nacimiento,
    talla_nacimiento,
    perimetro_craneal
  ) %>%

  tbl_strata(
    strata = madre_extranjera,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = macrosomia,
        missing = "no",
        percent = "column",

        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(
          edad ~ "Maternal age (years)",
          talla ~ "Maternal height (cm)",
          peso_inicial ~ "Initial maternal weight (kg)",
          imc_inicial ~ "Initial BMI (kg/m²)",
          ganancia_peso ~ "Gestational weight gain (kg)",
          consumo_tabaco ~ "Smoking during pregnancy",
          consumo_alcohol ~ "Alcohol consumption during pregnancy",

          edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
          tipo_parto ~ "Mode of delivery",

          peso_nacimiento ~ "Birth weight (kg)",
          talla_nacimiento ~ "Birth length (cm)",
          perimetro_craneal ~ "Head circumference (cm)"
        )

      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(
    "**Table 1. Baseline maternal, pregnancy and neonatal characteristics according to maternal origin (Spain vs Foreign) and LGA status. Values are presented as Median (Q1, Q3) or n (%).**"
  )

tabla1

save_table(tabla1, "Table1_Spain_vs_Foreign_LGA.html")
```  
  
### 4.1.1. Tabla suplementaria S1:  Baseline characteristics according to maternal origin
 
```{r}
tablaS1 <- df %>%
  select(
    madre_extranjera,

    ## Maternal characteristics
    edad,
    talla,
    peso_inicial,
    imc_inicial,
    ganancia_peso,
    consumo_tabaco,
    consumo_alcohol,

    ## Pregnancy
    edad_gestacional_nacimiento,
    tipo_parto,

    ## Neonatal
    peso_nacimiento,
    talla_nacimiento,
    perimetro_craneal,

    ## Obstetric complications
    diabetes_gestacional,
    hipertension_gestacional,
    preeclampsia,
    eclampsia,

    ## Prenatal care
    semanas_primera_visita,
    primera_visita_precoz,
    n_visitas_embarazo,
    visitas_esperadas,
    porcentaje_visitas,
    indice_kessner,
    indice_apncu,

    ## Healthcare utilisation
    n_visitas_atencion_primaria,
    n_visitas_hospitalarias

  ) %>%

  tbl_summary(
    by = madre_extranjera,
    missing = "no",
    percent = "column",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),
    digits = all_continuous() ~ 2,
    label = list(

      edad ~ "Maternal age (years)",
      talla ~ "Maternal height (cm)",
      peso_inicial ~ "Initial maternal weight (kg)",
      imc_inicial ~ "Initial BMI (kg/m²)",
      ganancia_peso ~ "Gestational weight gain (kg)",
      consumo_tabaco ~ "Smoking during pregnancy",
      consumo_alcohol ~ "Alcohol consumption during pregnancy",

      edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
      tipo_parto ~ "Mode of delivery",

      peso_nacimiento ~ "Birth weight (kg)",
      talla_nacimiento ~ "Birth length (cm)",
      perimetro_craneal ~ "Head circumference (cm)",

      diabetes_gestacional ~ "Gestational diabetes",
      hipertension_gestacional ~ "Gestational hypertension",
      preeclampsia ~ "Preeclampsia",
      eclampsia ~ "Eclampsia",

      semanas_primera_visita ~ "Gestational age at first prenatal visit (weeks)",
      primera_visita_precoz ~ "Early prenatal care initiation",
      n_visitas_embarazo ~ "Observed prenatal visits",
      visitas_esperadas ~ "Expected prenatal visits (ACOG)",
      porcentaje_visitas ~ "Observed/expected visits (%)",
      indice_kessner ~ "Kessner index",
      indice_apncu ~ "APNCU index",

      n_visitas_atencion_primaria ~ "Primary care visits",
      n_visitas_hospitalarias ~ "Hospital visits"
    )
  ) %>%

  tabla_estilo() %>%
  modify_table_body(
    ~ .x %>%

      dplyr::add_row(label = "**Maternal characteristics**", 
                     row_type = "label", .before = 1) %>%

      dplyr::add_row(label = "**Pregnancy**", row_type = "label", 
                     .before = which(.$variable == "edad_gestacional_nacimiento")) %>%

      dplyr::add_row(label = "**Neonatal characteristics**", row_type = "label",
                     .before = which(.$variable == "peso_nacimiento")) %>%

      dplyr::add_row(label = "**Maternal comorbidities**", row_type = "label",
                     .before = which(.$variable == "diabetes_gestacional")) %>%

      dplyr::add_row(label = "**Prenatal care adequacy**", row_type = "label",
                     .before = which(.$variable == "semanas_primera_visita")) %>%

      dplyr::add_row(label = "**Healthcare utilisation**", row_type = "label",
                     .before = which(.$variable == "n_visitas_atencion_primaria"))
  ) %>%

  modify_caption(
    "**Supplementary Table S1. Baseline maternal, pregnancy, neonatal characteristics, prenatal care adequacy and healthcare utilisation according to maternal origin. Values are presented as Median (Q1, Q3) or n (%).**"
)

tablaS1

save_table(tablaS1, "Supplementary_Table_S1_Maternal_Origin.html")
```  
  
## 4.2. Tabla 2: Baseline characteristics according to Global North vs Global South and LGA status
```{r}
tabla2 <- df %>%
  select(

    global_region,
    macrosomia,

    ## Maternal characteristics
    edad,
    talla,
    peso_inicial,
    imc_inicial,
    ganancia_peso,
    consumo_tabaco,
    consumo_alcohol,

    ## Pregnancy
    edad_gestacional_nacimiento,
    tipo_parto,

    ## Neonatal
    peso_nacimiento,
    talla_nacimiento,
    perimetro_craneal
  ) %>%

  tbl_strata(
    strata = global_region,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = macrosomia,
        missing = "no",
        percent = "column",

        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(
          edad ~ "Maternal age (years)",
          talla ~ "Maternal height (cm)",
          peso_inicial ~ "Initial maternal weight (kg)",
          imc_inicial ~ "Initial BMI (kg/m²)",
          ganancia_peso ~ "Gestational weight gain (kg)",
          consumo_tabaco ~ "Smoking during pregnancy",
          consumo_alcohol ~ "Alcohol consumption during pregnancy",

          edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
          tipo_parto ~ "Mode of delivery",

          peso_nacimiento ~ "Birth weight (kg)",
          talla_nacimiento ~ "Birth length (cm)",
          perimetro_craneal ~ "Head circumference (cm)"
        )

      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(
    "**Table 2. Baseline characteristics according to gloval region (Global North vs Global South) and LGA status. Values are presented as Median (Q1, Q3) or n (%).**" 
  )

tabla2

save_table(tabla2, "Table2_Global_Region_LGA.html")
```

## 4.3. Tabla 3: Maternal comorbidities associated with LGA

```{r}
tabla3 <- df %>%
  select(
    macrosomia,
    diabetes_gestacional,
    hipertension_gestacional,
    preeclampsia,
    eclampsia
  ) %>%

  tbl_summary(
    by = macrosomia,
    missing = "no",
    percent = "column",

    label = list(
      diabetes_gestacional ~ "Gestational diabetes",
      hipertension_gestacional ~ "Gestational hypertension",
      preeclampsia ~ "Preeclampsia",
      eclampsia ~ "Eclampsia"
    )

  ) %>%

  tabla_estilo() %>%

  modify_caption(
    "**Table 3. Maternal comorbidities and pregnancy complications associated with LGA.**"
  )

tabla3

save_table(tabla3, "Table3_Maternal_Diagnoses.html")
```

## 4.4. Tabla 4: Prenatal follow-up and healthcare utilisation
```{r}
tabla4 <- df %>%
  select(
    madre_extranjera ,
    macrosomia,
    
    ## Prenatal follow-up
    n_visitas_embarazo,
    
    ## Healthcare utilisation
    n_visitas_atencion_primaria,
    n_visitas_hospitalarias
  ) %>%

  tbl_strata(
    strata = madre_extranjera,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = macrosomia,
        missing = "no",
        percent = "column",
        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(

          n_visitas_embarazo ~ "Prenatal follow-up visits",
          
          n_visitas_atencion_primaria ~ "Primary care visits",

          n_visitas_hospitalarias ~ "Hospital visits"
        )
      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(

    "**Table 4. Prenatal follow-up and healthcare utilisation according to maternal origin and LGA status. Values are presented as Median (Q1, Q3).**"

  )

tabla4

save_table(tabla4, "Table4_Prenatal_Follow_Up.html")
```

## 4.5. Tabla 5: Adequacy of prenatal care according to neonatal macrosomia.
```{r}
tabla5 <- df %>%
  select(
    macrosomia,
    semanas_primera_visita,
    primera_visita_precoz,
    n_visitas_embarazo,
    visitas_esperadas,
    porcentaje_visitas,
    indice_kessner,
    indice_apncu
  ) %>%

  tbl_summary(
    by = macrosomia,
    missing = "no",
    percent = "column",

    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),

    digits = all_continuous() ~ 2,

    label = list(
      semanas_primera_visita ~ "Gestational age at first prenatal visit (weeks)",
      primera_visita_precoz ~ "Early prenatal care initiation",
      n_visitas_embarazo ~ "Observed prenatal visits",
      visitas_esperadas ~ "Expected prenatal visits (ACOG)",
      porcentaje_visitas ~ "Observed/expected visits (%)",
      indice_kessner ~ "Kessner index",
      indice_apncu ~ "APNCU index"
    )
  ) %>%

  tabla_estilo() %>%
  modify_table_body(
    ~ .x %>%
      dplyr::add_row(
        label = "**Timing of prenatal care**",row_type = "label",
        .before = which(.$variable == "semanas_primera_visita")
      ) %>%
      dplyr::add_row(
        label = "**Adequacy of prenatal care**",row_type = "label",
        .before = which(.$variable == "n_visitas_embarazo")
      )
    ) %>%
  
  modify_caption(
    "**Table 5. Adequacy of prenatal care according to neonatal macrosomia. Values are presented as Median (Q1, Q3) or n (%).**"
  )

tabla5

save_table(tabla5, "Table5_Prenatal_Care_Adequacy.html")
```  
  
---  
  
# Tablas alternativas   
Estas tablas descriptivas se construyen empleando las entidades del diagrama entidad-relación en vez de con el dataset de análisis.  
  
# 5. Construcción del dataset de análisis

## 5.1. Variables derivadas

```{r}
df <- hijo %>%
  left_join(embarazo, by = "id_embarazo") %>%
  left_join(madre, by = "id_madre") %>%
  left_join(uso_servicio, by = "id_embarazo") %>%
  mutate(

    # Outcome
    LGA = factor(
      if_else(peso_nacimiento >= 4000, "LGA", "No LGA"),
      levels = c("No LGA", "LGA")
    ),

    # Maternal origin
    foreign = factor(
      if_else(nacionalidad == "ESPAÑA", "Spain", "Foreign"),
      levels = c("Spain", "Foreign")
    ),

    global_region = case_when(
  
        nacionalidad %in% c(
          "ESPAÑA",
      
          # Europa
          "RUMANIA",
          "BULGARIA",
          "FRANCIA",
          "POLONIA",
          "UCRANIA",
          "MOLDAVIA",
          "PORTUGAL",
          "ITALIA",
          "FEDERACION RUSA",
          "BELGICA",
          "ALEMANIA",
          "PAISES BAJOS",
          "ARMENIA",
          "BIELORRUSIA",
          "HUNGRIA",
          "LITUANIA",
          "BOSNIA-HERZEGOVINA",
          "GEORGIA",
          "REPUBLICA CHECA",
          "ALBANIA",
          "TURQUIA",
          "ESLOVAQUIA",
          "LATVIA-LETONIA",
          "GRECIA",
          "SERBIA",
          "AUSTRIA",
          "FINLANDIA",
          "R.U.GRAN BRETAÑA E IRL N.",
      
          # Norteamérica (alto ingreso)
          "CANADA",
          "ESTADOS UNIDOS DE AMERICA"
        ) ~ "Global North",
      
        TRUE ~ "Global South"
      
      ),
      
      global_region = factor(
        global_region,
        levels = c("Global North", "Global South")
      )
  )


df <- df %>%
  mutate(

    consumo_tabaco = factor(
      consumo_tabaco,
      levels = c(0,1),
      labels = c("No","Yes")
    ),

    consumo_alcohol = factor(
      consumo_alcohol,
      levels = c(0,1),
      labels = c("No","Yes")
    ),

    muerte_neonatal = factor(
      muerte_neonatal,
      levels = c(0,1),
      labels = c("No","Yes")
    )

  )
```

## 5.2. Función para establecer el estilo de la tabla  
```{r}
tabla_estilo <- function(tabla){

  tabla %>%
    add_overall(last = FALSE) %>%
    add_n() %>%
    add_p(
      test = list(
        all_continuous() ~ "wilcox.test",
        all_categorical() ~ "chisq.test"
      )
    ) %>%
    bold_labels() %>%
    bold_p(t = 0.05) %>%
    italicize_levels()

}

tabla_estilo_strata <- function(tabla){

  tabla %>%
    add_n() %>%
    add_p(
      test = list(
        all_continuous() ~ "wilcox.test",
        all_categorical() ~ "chisq.test"
      )
    ) %>%
    bold_labels() %>%
    bold_p(t = 0.05) %>%
    italicize_levels()

}
```  
  
## 5.3. Función para guardar las tablas  
```{r}
save_table <- function(tbl, filename){

  tbl %>%
    as_gt() %>%
    gt::gtsave(
      paste0(
        "Y:/PROYECTOS/2024 Salud perinatal (Luis-Aída-Sol)/Desarrollo/Datos/Results/tablas/",
        filename
      )
    )

}
```  
  
## 6. Tablas  
  
## 6.1. Tabla 1: Overall characteristics according to LGA.
 
```{r}
tabla1 <- df %>%
  select(
    LGA,
    
      ## Maternal characteristics
      edad,
      foreign,
      talla,
      peso_inicial,
      imc_inicial,
      ganancia_peso,
      consumo_tabaco,
      consumo_alcohol,
    
      ## Pregnancy
      tipo_parto,
    
      ## Neonatal
      peso_nacimiento,
      talla_nacimiento,
      perimetro_craneal,
      edad_gestacional_nacimiento,
      muerte_neonatal
    ) %>%
  
  tbl_summary(
    by = LGA,
    missing = "no",
    percent = "column",
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})"
    ),
  
    digits = all_continuous() ~ 2,
  
    label = list(
      edad ~ "Maternal age (years)",
      foreign ~ "Maternal origin",
      talla ~ "Maternal height (cm)",
      peso_inicial ~ "Initial maternal weight (kg)",
      imc_inicial ~ "Initial BMI (kg/m²)",
      ganancia_peso ~ "Gestational weight gain (kg)",
      consumo_tabaco ~ "Smoking during pregnancy",
      consumo_alcohol ~ "Alcohol consumption during pregnancy",
      tipo_parto ~ "Mode of delivery",
      peso_nacimiento ~ "Birth weight (g)",
      talla_nacimiento ~ "Birth length (cm)",
      perimetro_craneal ~ "Head circumference (cm)",
      edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
      muerte_neonatal ~ "Neonatal death"
    )
  ) %>%
  
  tabla_estilo() %>%
  
  modify_caption(
    "**Table 1. Overall maternal, pregnancy and neonatal characteristics according to LGA status. Values are presented as Median (Q1, Q3) or n (%).**"
  )

tabla1

save_table(tabla1, "Table1_LGA.html")
```
  
## 6.2. Tabla 2: Baseline characteristics according to maternal origin (Spain vs Foreign) and LGA status.

```{r}
tabla2 <- df %>%
  select(

    foreign,
    LGA,

    ## Maternal characteristics
    edad,
    talla,
    peso_inicial,
    imc_inicial,
    ganancia_peso,
    consumo_tabaco,
    consumo_alcohol,

    ## Obstetric history
    embarazos_anteriores,
    nacimientos_anteriores,
    abortos_anteriores,

    ## Pregnancy
    edad_gestacional_nacimiento,
    tipo_parto,

    ## Neonatal
    peso_nacimiento,
    talla_nacimiento,
    perimetro_craneal,
    muerte_neonatal

  ) %>%

  tbl_strata(
    strata = foreign,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = LGA,
        missing = "no",
        percent = "column",

        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(
          edad ~ "Maternal age (years)",
          talla ~ "Maternal height (cm)",
          peso_inicial ~ "Initial maternal weight (kg)",
          imc_inicial ~ "Initial BMI (kg/m²)",
          ganancia_peso ~ "Gestational weight gain (kg)",
          consumo_tabaco ~ "Smoking during pregnancy",
          consumo_alcohol ~ "Alcohol consumption during pregnancy",

          embarazos_anteriores ~ "Previous pregnancies",
          nacimientos_anteriores ~ "Previous live births",
          abortos_anteriores ~ "Previous miscarriages",

          edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
          tipo_parto ~ "Mode of delivery",

          peso_nacimiento ~ "Birth weight (g)",
          talla_nacimiento ~ "Birth length (cm)",
          perimetro_craneal ~ "Head circumference (cm)",
          muerte_neonatal ~ "Neonatal death"
        )

      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(
    "**Table 2. Baseline maternal, pregnancy and neonatal characteristics according to maternal origin (Spain vs Foreign) and LGA status. Values are presented as Median (Q1, Q3) or n (%).**"
  )

tabla2

save_table(tabla2, "Table2_Spain_vs_Foreign_LGA.html")
```

## 6.3. Tabla 3: Baseline characteristics according to Global North vs Global South and LGA status
```{r}
tabla3 <- df %>%
  select(

    global_region,
    LGA,

    ## Maternal characteristics
    edad,
    talla,
    peso_inicial,
    imc_inicial,
    ganancia_peso,
    consumo_tabaco,
    consumo_alcohol,

    ## Obstetric history
    embarazos_anteriores,
    nacimientos_anteriores,
    abortos_anteriores,

    ## Pregnancy
    edad_gestacional_nacimiento,
    tipo_parto,

    ## Neonatal
    peso_nacimiento,
    talla_nacimiento,
    perimetro_craneal,
    muerte_neonatal

  ) %>%

  tbl_strata(
    strata = global_region,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = LGA,
        missing = "no",
        percent = "column",

        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(
          edad ~ "Maternal age (years)",
          talla ~ "Maternal height (cm)",
          peso_inicial ~ "Initial maternal weight (kg)",
          imc_inicial ~ "Initial BMI (kg/m²)",
          ganancia_peso ~ "Gestational weight gain (kg)",
          consumo_tabaco ~ "Smoking during pregnancy",
          consumo_alcohol ~ "Alcohol consumption during pregnancy",

          embarazos_anteriores ~ "Previous pregnancies",
          nacimientos_anteriores ~ "Previous live births",
          abortos_anteriores ~ "Previous miscarriages",

          edad_gestacional_nacimiento ~ "Gestational age at birth (weeks)",
          tipo_parto ~ "Mode of delivery",

          peso_nacimiento ~ "Birth weight (g)",
          talla_nacimiento ~ "Birth length (cm)",
          perimetro_craneal ~ "Head circumference (cm)",
          muerte_neonatal ~ "Neonatal death"
        )

      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(
    "**Table 3. Baseline characteristics according to maternal origin (Global North vs Global South) and LGA status. Values are presented as Median (Q1, Q3) or n (%).**" 
  )

tabla3

save_table(tabla3, "Table3_Global_Region_LGA.html")
```

## 6.4. Tabla 4: Maternal comorbidities associated with LGA

```{r}
diag_madre <- diagnostico %>%
  filter(tipo_paciente == "Madre") %>%
  mutate(
    txt = str_to_upper(diag_st),

    diabetes =
      str_detect(txt, "DIABETES") &
      !str_detect(txt, "GESTACIONAL"),

    gestational_diabetes =
      str_detect(txt, "DIABETES GESTACIONAL"),

    hypertension =
      str_detect(txt, "HIPERTENS") &
      !str_detect(txt, "GESTACIONAL"),

    gestational_hypertension =
      str_detect(txt, "HIPERTENSION GESTACIONAL"),

    preeclampsia =
      str_detect(txt, "PREECLAMP"),

    eclampsia =
      str_detect(txt, "ECLAMP")
  ) %>%

  group_by(id_paciente) %>%

  summarise(
    diabetes = any(diabetes),
    gestational_diabetes = any(gestational_diabetes),
    hypertension = any(hypertension),
    gestational_hypertension = any(gestational_hypertension),
    preeclampsia = any(preeclampsia),
    eclampsia = any(eclampsia),

    .groups = "drop"
  )

tabla4A <- df %>%

  left_join(
    diag_madre,
    by = c("id_madre" = "id_paciente")
  ) %>%

  mutate(
    across(
      c(
        diabetes,
        gestational_diabetes,
        hypertension,
        gestational_hypertension,
        preeclampsia,
        eclampsia
      ),
      ~replace_na(.x, FALSE)
    )
  ) %>%

  select(
    LGA,

    diabetes,
    gestational_diabetes,

    hypertension,
    gestational_hypertension,

    preeclampsia,
    eclampsia

  ) %>%

  tbl_summary(
    by = LGA,
    missing = "no",
    percent = "column",

    label = list(
      diabetes ~ "Pregestational diabetes",
      gestational_diabetes ~ "Gestational diabetes",
      hypertension ~ "Chronic hypertension",
      gestational_hypertension ~ "Gestational hypertension",
      preeclampsia ~ "Preeclampsia",
      eclampsia ~ "Eclampsia"
    )

  ) %>%

  tabla_estilo() %>%

  modify_caption(
    "**Table 4A. Maternal comorbidities and pregnancy complications associated with LGA.**"
  )

tabla4A

save_table(tabla4A, "Table4A_Maternal_Diagnoses.html")
```

## 6.6. Tabla 5: Prenatal follow-up and healthcare utilisation
```{r}
tabla5 <- df %>%
  select(
    foreign,
    LGA,
    
    ## Prenatal follow-up
    n_visitas_embarazo,

    ## Healthcare utilisation
    n_visitas_atencion_primaria,
    n_visitas_hospitalarias

  ) %>%

  tbl_strata(
    strata = foreign,
    .tbl_fun = ~ .x %>%

      tbl_summary(
        by = LGA,
        missing = "no",
        percent = "column",
        statistic = list(
          all_continuous() ~ "{median} ({p25}, {p75})"
        ),

        digits = all_continuous() ~ 2,

        label = list(

          n_visitas_embarazo ~ "Prenatal follow-up visits",
          
          n_visitas_atencion_primaria ~ "Primary care visits",

          n_visitas_hospitalarias ~ "Hospital visits"
        )
      ) %>%

      tabla_estilo_strata()

  ) %>%

  modify_caption(

    "**Table 5. Prenatal follow-up and healthcare utilisation according to maternal origin and LGA status. Values are presented as Median (Q1, Q3).**"

  )

tabla5

save_table(tabla5, "Table5_Prenatal_Follow_Up.html")
```

